\textbf{NBA Party}

### 2017-07-06 init_model.rnw
Now only betting once per game!

require(hangover)
require(data.table)
require(ggplot2)
require(magrittr)
require(sampling)
require(stargazer)
set.seed(4052017)

@@

##################################################
### Table for uncorrelated next day opponent
##################################################

@@

@@

pdf('writing/why_drinks_or_population_dont_work.pdf', width = 16, height = 8.5)
ticks = 1:nrow(tmp)
plot(x = ticks, 
     y = tmp$log.nmusicians,
     main = 'Why drinking establishments or population don\'t work as a proxy for nightlife',
     ylab = 'Log sum of Variable',
     xlab = '', axes = F, pch = 1, ylim = c(0, max(tmp$log.population)), type = 'l')
points(x = ticks, y = tmp$log.ndrinks, pch = 16, col = 'purple', type = 'l')
points(x = ticks, y = tmp$log.population, pch = 18, col = 'orange', type = 'l')
### Houston has large population, but isn't necessarily known for nightlife.
segments(x0 = which(tmp$team == 'houston rockets'), 
         x1 = which(tmp$team == 'houston rockets'),
         y0 = -1,
         y1 = tmp[team == 'houston rockets', log.population], col = 'gray70', lty = 'dashed')
segments(x0 = which(tmp$team == 'houston rockets'), 
         x1 = length(ticks),
         y0 = tmp[team == 'houston rockets', log.population],
         y1 = tmp[team == 'houston rockets', log.population], col = 'gray70', lty = 'dashed')
abline(h = tmp[team == 'houston rockets', log.ndrinks], col = 'gray70', lty = 'dashed')
### New orleans is infamous for party, but Population places it at the bottom of the index.
segments(x0 = which(tmp$team == 'new orleans hornets'),
         x1 = which(tmp$team == 'new orleans hornets'), 
         y0 = -1, 
         y1 = tmp[team == 'new orleans hornets', log.population],col = 'gray70', lty = 'dashed')
segments(x0 = 0, x1 = which(tmp$team == 'new orleans hornets'),
         y0 = tmp[team == 'new orleans hornets', log.population],
         y1 = tmp[team == 'new orleans hornets', log.population], col = 'gray70', lty = 'dashed')
axis(side = 1, at = ticks, labels = gsub('([a-z]{4,}) ', '\\1\n', tmp$team) %>%
                               gsub('([a-z]{6})[a-z]{1,}', '\\1', .), las = 2, xpd = NA)
axis(side = 2)
legend(x = 1.5, y = 12.5, legend = c('Musicians and sound recording studios',
                                 'Drinking establishments and liquor stores',
                                 'Population'), pch = c(1, 16, 18), col = c('black', 'purple', 'orange'))
text(x = 3.6, y = 9.5, labels = paste("Correlation (alcohol):       ", 
                                       cor(tmp$log.nmusicians, tmp$log.ndrinks) %>% round(., 3)))
text(x = 3.6, y = 10, labels = paste("Correlation (population): ", 
                                       cor(tmp$log.nmusicians, tmp$log.population) %>% round(., 3)))
dev.off()

@@

pdf('writing/mlb_bets_by_season.pdf')
par(mfrow = c(3,2))
b <- lapply(2012:2016, function(yr) {
    m <- glm(I(team.score > opponent.score) ~ party + odds,
             data = mlb.lines[year(date) < yr], family = 'binomial')
    p <- predict(m, newdata = mlb.lines[year(date) == yr], type = 'response')
    idx <- which(p < .45)
    p1 <- mapply(Predict, prediction = p[idx], 
                 vegas = mlb.lines[year(date) == yr, odds][idx],
                 MoreArgs=list(threshold = 0.01))
    b <- mapply(Bet, our.prediction = p1,
                actual.outcome = mlb.lines[year(date) == yr, team.score > opponent.score][idx],
                odds = mlb.lines[year(date) == yr, odds][idx], SIMPLIFY = T)
    cat(paste0(sum(b, na.rm = T), '\n'))
    ###
#    pdf(paste0('mlb_bets_', yr, '.pdf'))
    betting.days <- mlb.lines[year(date) == yr][idx, date]
    running.prof <- cumsum(ifelse(is.na(b), 0, b))
    plot(x = 1:length(b), y = running.prof, 
         main = paste0('Profits over the ', yr, ' season\n',
                       'Training on 2011-', yr-1, ' data'),
         xlab = 'Game Number', ylab = 'Profit', type = 'l')
    abline(a = 0, b = 0, lty = 3)
    points(x = which.min(running.prof), y = min(running.prof), type = 'p')
    text(x = which.min(running.prof), y = min(running.prof), 
         labels = round(min(running.prof)), pos = 4)
    return(b)
#    dev.off()
})
dev.off()

### Meet the spread for both NBA and MLB
pdf('writing/meet_the_spread.pdf')
par(mfcol = c(2, 1))
tmp <- nba.lines[, list(meet.spread = mean(outcome == 'W')), 
          by = list(season, team)]
hist(tmp$meet.spread, prob = T, main = paste('How often do teams meet the spread? (NBA)', 
                                             'By season-team, 2010-11 through 2016-17',
                                             sep = '\n'),
     xlab = 'Meet the Spread', breaks = 25)
lines(density(tmp$meet.spread, adjust = 2), col = 'orange', lty = 'dashed')
### There should be 'no information' left over after conditioning on odds.
noinf <- lm(I(team.score > opponent.score) ~ odds, data = mlb.lines)
preds <- fitted(noinf)
mlb.lines[, fitted := fitted(noinf)]
mlb.lines[, resid  := resid(noinf)]
mlb.lines[, list(avg.residual = mean(resid)), by = round(fitted, 3)] %>% 
    plot(., xlab = 'Fitted Probability', ylab = 'Average Residual', 
         main = paste('After conditioning on bookkeeper\'s odds (MLB),',
                      'successful prediction is a coin-flip', sep = '\n'))
abline(a = 0, b = 0, col = 'orange', lty = 'dashed')
dev.off()

@@

m <- glm(I(team.score > opponent.score) ~ party + odds,
         data = mlb.lines[year(date) < 2016], family = 'binomial')
p <- predict(m, newdata = mlb.lines[year(date) == 2016], type = 'response')
idx <- which(p < .45)
p1 <- mapply(Predict, prediction = p[idx], 
             vegas = mlb.lines[year(date) == 2016, odds][idx],
             MoreArgs=list(threshold = 0.01))
b <- mapply(Bet, our.prediction = p1,
                actual.outcome = mlb.lines[year(date) == 2016, team.score > opponent.score][idx],
            odds = mlb.lines[year(date) == 2016, odds][idx], SIMPLIFY = T)

### the following plot shows sufficient deviation from house odds to motivate
### an optimized betting strategy.
hist(p- mlb.lines[year(date) == 2016, odds])

tholds <- seq(.01, .1, by = .005)
tmp <- sapply(tholds, function(t) {
    p1 <- mapply(Predict, prediction = p[idx], 
                 vegas = mlb.lines[year(date) == 2016, odds][idx],
                 MoreArgs=list(threshold = t))
    b <- mapply(Bet, our.prediction = p1,
                actual.outcome = mlb.lines[year(date) == 2016, team.score > opponent.score][idx],
                odds = mlb.lines[year(date) == 2016, odds][idx], SIMPLIFY = T)
    return(c(sum(b, na.rm = T) / sum(!is.na(b)), sum(!is.na(b)))) # Avg profit per game.
})

>>>>

require(hangover)

require(data.table)
require(ggplot2)
require(magrittr)
require(sampling)
require(stargazer)
set.seed(4052017)

@@

##################################################
### Table for (un)-correlated next day opponent
##################################################

@@

## tmp <- nba.lines[, list(meet.spread = mean(outcome == 'W'), .N), by = list(last.game.loc, team)][order(N)]
## hist(tmp$meet.spread, probability = T)
## lines(density(tmp$meet.spread, adjust = 1))

@@

pdf('writing/why_drinks_or_population_dont_work.pdf', width = 16, height = 8.5)
ticks = 1:nrow(tmp)
plot(x = ticks, 
     y = tmp$log.nmusicians,
     main = 'Why drinking establishments or population don\'t work as a proxy for nightlife',
     ylab = 'Log sum of Variable',
     xlab = '', axes = F, pch = 1, ylim = c(0, max(tmp$log.population)), type = 'l')
points(x = ticks, y = tmp$log.ndrinks, pch = 16, col = 'purple', type = 'l')
points(x = ticks, y = tmp$log.population, pch = 18, col = 'orange', type = 'l')
### Houston has large population, but isn't necessarily known for nightlife.
segments(x0 = which(tmp$team == 'houston rockets'), 
         x1 = which(tmp$team == 'houston rockets'),
         y0 = -1,
         y1 = tmp[team == 'houston rockets', log.population], col = 'gray70', lty = 'dashed')
segments(x0 = which(tmp$team == 'houston rockets'), 
         x1 = length(ticks),
         y0 = tmp[team == 'houston rockets', log.population],
         y1 = tmp[team == 'houston rockets', log.population], col = 'gray70', lty = 'dashed')
abline(h = tmp[team == 'houston rockets', log.ndrinks], col = 'gray70', lty = 'dashed')
### New orleans is infamous for party, but Population places it at the bottom of the index.
segments(x0 = which(tmp$team == 'new orleans hornets'),
         x1 = which(tmp$team == 'new orleans hornets'), 
         y0 = -1, 
         y1 = tmp[team == 'new orleans hornets', log.population],col = 'gray70', lty = 'dashed')
segments(x0 = 0, x1 = which(tmp$team == 'new orleans hornets'),
         y0 = tmp[team == 'new orleans hornets', log.population],
         y1 = tmp[team == 'new orleans hornets', log.population], col = 'gray70', lty = 'dashed')
axis(side = 1, at = ticks, labels = gsub('([a-z]{4,}) ', '\\1\n', tmp$team) %>%
                               gsub('([a-z]{6})[a-z]{1,}', '\\1', .), las = 2, xpd = NA)
axis(side = 2)
legend(x = 1.5, y = 12.5, legend = c('Musicians and sound recording studios',
                                 'Drinking establishments and liquor stores',
                                 'Population'), pch = c(1, 16, 18), col = c('black', 'purple', 'orange'))
text(x = 3.6, y = 9.5, labels = paste("Correlation (alcohol):       ", 
                                       cor(tmp$log.nmusicians, tmp$log.ndrinks) %>% round(., 3)))
text(x = 3.6, y = 10, labels = paste("Correlation (population): ", 
                                       cor(tmp$log.nmusicians, tmp$log.population) %>% round(., 3)))
points(x = c(which(tmp$team == 'new orleans hornets'),
             which(tmp$team == 'houston rockets'),
             which(tmp$team == 'houston rockets')),
       y = c(tmp[team == 'new orleans hornets', log.population],
             tmp[team == 'houston rockets', log.population],
             tmp[team == 'houston rockets', log.ndrinks]),
       type = 'p', col = 'red')
dev.off()

@@

pdf('writing/mlb_bets_by_season.pdf')
par(mfrow = c(3,2))
b <- lapply(2012:2016, function(yr) {
    m <- glm(I(team.score > opponent.score) ~ party + odds,
             data = mlb.lines[season < yr], family = 'binomial')
    p <- predict(m, newdata = mlb.lines[year(date) == yr], type = 'response')
    ### Subset data to relevant year we wish to predict on; make predictions.
    tmp <- mlb.lines[season == yr, .(matchup, date, team, opponent, 
                                     team.score, opponent.score, odds)]
    tmp[, fitted := predict(m, newdata = mlb.lines[year(date) == yr], type = 'response')]
    ### Realize that we can only make one prediction per game. So take the stronger bet.
    setkey(tmp, date, matchup)
    tmp <- tmp[, .SD[which.max(abs(fitted - .5))], by = list(matchup, date)]
    tmp <- mlb.lines[tmp, on = intersect(names(tmp), names(mlb.lines))]
    ### Translate fitted probabilities into 0-1 predictions if our odds differ by threshold amount.
    tmp[, prediction := mapply(hangover::Predict, prediction = tmp$fitted, vegas = tmp$odds, 
                               MoreArgs = list(threshold = 0.02))] # 0.02 was chosen ad-hoc.
    tmp <- tmp[!is.na(prediction)]
    ### Translate prediction into a monetary gain or loss based on odds and actual outcome.
    tmp[, bet := mapply(hangover::Bet, our.prediction = prediction, 
                        actual.outcome = team.score > opponent.score,
                        odds = odds)]
    ### Plot the cumulative sum of profits over the course of the season.
    tmp[, running.prof := cumsum(bet)]    
    cat(paste0(sum(tmp$bet), '\n'))
    plot(x = 1:nrow(tmp), y = tmp$running.prof, 
         main = paste0('Profit over the ', yr, ' season: $', 
                       format(round(sum(tmp$bet), digits = -1), big.mark = ','), '\n',
                       'Training on 2011-', yr-1, ' data'),
         xlab = 'Game Number', ylab = 'Profit', type = 'l')
    abline(a = 0, b = 0, lty = 3)
    points(x = which.min(tmp$running.prof), y = min(tmp$running.prof), type = 'p')
    text(x = which.min(tmp$running.prof), y = min(tmp$running.prof), 
         labels = round(min(tmp$running.prof)), pos = 4)
    return(tmp[, bet])
})
dev.off()

### Meet the spread for both NBA and MLB
pdf('writing/meet_the_spread.pdf')
par(mfcol = c(2, 1))
tmp <- nba.lines[, list(meet.spread = mean(outcome == 'W')), 
          by = list(season, team)]
hist(tmp$meet.spread, prob = T, main = paste('How often do teams meet the spread? (NBA)', 
                                             'By season-team, 2010-11 through 2016-17',
                                             sep = '\n'),
     xlab = 'Meet the Spread', breaks = 25)
lines(density(tmp$meet.spread, adjust = 2), col = 'orange', lty = 'dashed', lwd = 3)
### There should be 'no information' left over after conditioning on odds.
noinf <- lm(I(team.score > opponent.score) ~ odds, data = mlb.lines)
preds <- fitted(noinf)
mlb.lines[, fitted := fitted(noinf)]
mlb.lines[, resid  := resid(noinf)]
tmp <- mlb.lines[, .(avg.residual = mean(resid)), by = .(fitted = round(fitted, 3))]
plot(x = tmp$fitted, y = tmp$avg.residual, 
     xlab = 'Fitted Probability', ylab = 'Average Residual', 
     main = paste('After conditioning on bookkeeper\'s odds (MLB),',
                  'successful prediction is a coin-flip', sep = '\n'))
abline(lm(avg.residual ~ fitted, tmp), col = 'orange', lty = 'dashed', lwd = 3)
dev.off()

@@

m <- glm(I(team.score > opponent.score) ~ party + odds,
         data = mlb.lines[season < 2016], family = 'binomial')
p <- predict(m, newdata = mlb.lines[season == 2016], type = 'response')
idx <- which(p < .45)
p1 <- mapply(Predict, prediction = p[idx], 
             vegas = mlb.lines[season == 2016, odds][idx],
             MoreArgs=list(threshold = 0.01))
b <- mapply(Bet, our.prediction = p1,
                actual.outcome = mlb.lines[season == 2016, team.score > opponent.score][idx],
            odds = mlb.lines[season == 2016, odds][idx], SIMPLIFY = T)

### the following plot shows sufficient deviation from house odds to motivate
### an optimized betting strategy.
hist(p- mlb.lines[season == 2016, odds])

tholds <- seq(.01, .1, by = .005)
tmp <- sapply(tholds, function(t) {
    p1 <- mapply(Predict, prediction = p[idx], 
                 vegas = mlb.lines[season == 2016, odds][idx],
                 MoreArgs=list(threshold = t))
    b <- mapply(Bet, our.prediction = p1,
                actual.outcome = mlb.lines[season == 2016, team.score > opponent.score][idx],
                odds = mlb.lines[season == 2016, odds][idx], SIMPLIFY = T)
    return(c(sum(b, na.rm = T) / sum(!is.na(b)), sum(!is.na(b)))) # Avg profit per game.
})
